\pkg{Rgbp}: An R Package for Gaussian, Poisson, and Binomial Hierarchical Modeling

### 2013-08-03 jss/article.tex
changes

For this purpose, data were gathered from 31 hospitals in NY composed of the number of deaths ($z$) within a month of CABG surgeries and total number of patients ($n$) receiving CABG surgeries in each hospital. Using these data, \pkg{Rgbp} provides point and interval estimates of the true death ratio for each hospital so that one can choose the most reliable hospital. The data can be loaded into R using the following code:
\begin{CodeChunk}
\begin{CodeInput}
R> z <- c( 3, 2, 5, 11, 9, 12, 12, 4, 10, 13, 14, 7, 12, 11, 13, 22, 15, 
          11, 14, 11, 16, 14, 9, 15, 13, 35, 26, 25, 20, 35, 27)
R> n <- c(67, 68, 210, 256, 269, 274, 278, 295, 347, 349, 358, 396, 431,
         441, 477, 484, 494, 501, 505, 540, 563, 593, 602, 629, 636, 729,
         849, 914, 940, 1193, 1340)
\end{CodeInput}
\end{CodeChunk}
or
\begin{CodeChunk}
\begin{CodeInput}
R> data(hospitals)
\end{CodeInput}
\end{CodeChunk}

@@

Next, \code{gbp} will help us fit the Poisson hierarchical model with the Gamma conjugate prior distribution on the true death rate $\lambda_{j}$ whose mean is 0.020 ($\lambda_{0}=0.020$) as described in the section 2.2. For reference, the number of regression coefficients ($m$) is 0 because we do not need to estimate the prior mean via log-linear regression (see section 3.2).
\begin{CodeChunk}
\begin{CodeInput}
R> data(hospital)
R> z <- hospital$d
R> n <- hospital$n
R> p <- gbp(z, n, mean.PriorDist = 0.02, model = "pr")
R> p
\end{CodeInput}
\begin{CodeOutput}
Summary for each unit (sorted by n):

@@

Due to the fact that the estimation procedures in \pkg{Rgbp} rely on differentiation the package is extremely quick in fitting the available models. This makes the package ideal to be used in simulation studies where a hierarchical model needs to be fitted at every iteration and where running a full Bayesian model (via MCMC) would be computationally impractical.

\section[acknowledgments]{Acknowledgments}
The authors would like to thank Phil Everson, Cindy Christiansen, and the 2012 class of Stat 324r: Parametric Statistical Inference and Modeling for their valuable inputs.

>>>>

For this purpose, data were gathered from 31 hospitals in NY composed of the number of deaths ($z$) within a month of CABG surgeries and total number of patients ($n$) receiving CABG surgeries in each hospital. Using these data, \pkg{Rgbp} provides point and interval estimates of the true death ratio for each hospital so that one can choose the most reliable hospital. The data can be loaded into R using the following code:
\begin{CodeChunk}
\begin{CodeInput}
R> z <- c( 3, 2, 5, 11, 9, 12, 12, 4, 10, 13, 14, 7, 12, 11, 13, 22, 15, 
          11, 14, 11, 16, 14, 9, 15, 13, 35, 26, 25, 20, 35, 27)
R> n <- c(67, 68, 210, 256, 269, 274, 278, 295, 347, 349, 358, 396, 431,
         441, 477, 484, 494, 501, 505, 540, 563, 593, 602, 629, 636, 729,
         849, 914, 940, 1193, 1340)
\end{CodeInput}
\end{CodeChunk}
or
\begin{CodeChunk}
\begin{CodeInput}
R> data(hospitals)
R> z <- hospitals$d
R> n <- hospitals$n
\end{CodeInput}
\end{CodeChunk}

@@

Next, \code{gbp} will help us fit the Poisson hierarchical model with the Gamma conjugate prior distribution on the true death rate $\lambda_{j}$ whose mean is 0.020 ($\lambda_{0}=0.020$) as described in the section 2.2. For reference, the number of regression coefficients ($m$) is 0 because we do not need to estimate the prior mean via log-linear regression (see section 3.2).
\begin{CodeChunk}
\begin{CodeInput}
R> p <- gbp(z, n, mean.PriorDist = 0.02, model = "pr")
R> p
\end{CodeInput}
\begin{CodeOutput}
Summary for each unit (sorted by n):

@@

Due to the fact that the estimation procedures in \pkg{Rgbp} rely on derivatives, the package is extremely quick in fitting the available models. This makes the package ideal to be used in simulation studies where a hierarchical model needs to be fitted at every iteration and where running a full Bayesian model (via MCMC) would be computationally impractical.

\section[acknowledgments]{Acknowledgments}
The authors thank Prof Phil Everson of Swarthmore and the 2012 class of Harvard's Stat 324r: Parametric Statistical Inference and Modeling for their valuable inputs.
